\documentclass{article}

% Use the Cactus ThornGuide style file
% (Automatically used from Cactus distribution, if you have a 
%  thorn without the Cactus Flesh download this from the Cactus
%  homepage at www.cactuscode.org)
\usepackage{../../../../doc/latex/cactus}

\begin{document}

\title{Nice}
\author{Frank L\"offler (knarf@cct.lsu.edu)}

\maketitle

% Do not delete next line
% START CACTUS THORNGUIDE

\begin{abstract}
Simple tool for using a (configurable) nice level for all processes.
\end{abstract}

\section{Comments}

This thorn tries to renice all processes of Cactus to a given nice level.\\
To renice a process means to change its nice level after the start. The nice
level is a way to control the priority of processes for cpu power. It is
always a good idea to give cpu-intensive jobs (like Cactus runs) a low
priority if you are using desktop machines, especially if other people are
using them too.

The nice level has a range of $-20$ to $19$. A higher number means less
priority. Mostly, normal processes will get $0$ and only the superuser
(root) can use negative values. Also only root can lower the nice level
of a process, so if you nice your Cactus Run to 19, you cannot renice it
later to an lower nice-level. The nice level of single
processes can be controled by the system commands $nice$ and/or $renice$.
In the case of Cactus and Lamd this is not possible at startup. Therefore
this thorn will set the nice level of your processes at CCTK\_PARAMCHECK
(in the beginning of the calculation).\\
This thorn has only one parameter:
\begin{verbatim}
  CCTK_INT Nice_nice
\end{verbatim}
which is by default 19.\\*[1cm]
Note to developers:\\
This thorn uses $getpriority()$ and $setpriority()$ and therefore needs
the includes $<$sys/time.h$>$, $<$errno.h$>$ and $<$sys/resource.h$>$.
If these are not provided on some platform, please tell me.
% Do not delete next line
% END CACTUS THORNGUIDE

\end{document}
